### 2016-01-15 3_AB_test/AB_test.Rmd
abtest done for now

---
title: "Online A / B test"
author: "Ming-Yu Liu"
date: "January 15, 2016"
output: 
  html_document: 
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Overview

A / B testing is a general methodology used online to when you want to test out a new feature. What you're doing is you're going show one set of features, the control set ( your existing feature ) to one user group and another set, your experiment set ( your new feature ) to another user group and test how did these users respond differently so that you can determine which set of your feature is better.

Despite its useful functionality, there are still places where A / B testing isn't as useful. For example :

1. Testing out completely new experiences. An obvious way for thinking about this is if you're dramatically changing the user's experience. The user might prefer the old way of doing things or they've curious about the new features and they simply test out everything ( sometimes called the novelty effect ).
2. A / B testing can't tell you if you're missing something. Meaning it can tell you if A performs better B or vice versa, but it can't tell you that if you use C, then it will actually perform better than the former two.
3. Tesing out products that people rarely buy. e.g. cars, apartments. It might be too long before the user actually decides to take actions after seeing the information and you might be unaware of the actual motivation.

# Quick Example

When launching a A / B test you need to ask yourself :

1. What is your hypothesis and what's the baseline for comparison. 
2. What's the confidence level that you wish to have in your result ( This will also affect the number of samples ). 
2. How many samples and time do you need in order to actually have your user adapt to the new experience.

So now, suppose you're running an educational platform and your hypothesis is : Will changing the "Start Now" button from orange to pink increase how many students explore the platform's courses. So in this case the metric that's use to evaluate the change's performance is the click through probability ( Unique visitors who click the button / Unique visitors to page ). Note that it is often times impractical to use metrices such as total number of students that completed the course as it often takes weeks or months before a student can do that.

Next we will jot down the hypothesis that we wish to test out, in our case the our null and alternative hypothesis would be :

**The null hypothesis, H0: ** The experimental and control groups have the same probability of completing a checkout ( clicking the button ). Or equivalent to saying that the differences of the two groups' probability is 0.   
**The alternative hypothesis, H1: ** The two groups have different probability of completing a checkout.

Now that we've defined our hypothesis, the first question that comes into mind is how many tests do we need to run, or in a sense how long should the test last in order for us to make our decisions. To do that we can use a power analysis for two independent samples, which can be calculated using the `power.prop.test` function. Before doing that there are some basic parameters that we should be aware of :

1. **Significance Level: ** Denoted by $\alpha$, this threshold governs the chance of false positive. A significance level of 0.05 means that there's a 5 percent chance of false positive.
2. **Statistical power: ** This stands for $1 - \beta$ ( or so called sensitivity ), where $\beta$ represents the probability that you'll get a false negative. So a statistical power of 0.8 means that if there's is in fact a change, there's 80 percent that we'll detect it, which is equivalent to saying that there will be a 20 percent false negative.

The effect of picking a significance level of 0.05 and power of 0.8 means that we are 4 times more likely to obtain a false negative than a false positive. For A / B testing, we’re generally more concerned about getting a false positive : Making a change that doesn’t actually improve things than we are about not making a change at all, which is why we accept a greater likelihood of a false negative. The rule of thumb for $\alpha$ and $1 - \beta$ is 0.5 and 0.8 respectively.

Now suppose that our current baseline is 0.1 ( there's a 10 percent chance that people who saw the button will click it ). And we wish to detect a change of two percent in the click through rate ( This is already consider quite high for online experiment ).

Parameters :

- `baseline` Your current baseline solution. 
- `delta` Minimum detectable change, smallest effect that will be detected (1-β)% of the time. This parameter can also be referred to as the practical significance boundary.
- `power` Percent of the time the minimum effect size will be detected, assuming it exists.
- `sig_level` Percent of the time a difference will be detected, assuming one does NOT exist.

```{r}

# parameters
baseline  <- 0.1
delta 	  <- 0.02
power 	  <- 0.8
sig_level <- 0.05

result <- power.prop.test( p1 = baseline, p2 = baseline + delta, 
				 		   power = power, sig.level = sig_level,
				 		   alternative = "two.sided" )
result

```

The result shows that we need at least `r round(result$n)` sample size for each scenario to detect if there will actually be a 2 percent more-than-baseline click through probability.

How these parameters affect the sample size you need :

- `baseline` The higher the baseline click through probability ( but still less than 0.5 ), the larger the sample size you'll need. Since the probability is related to the standard deviation, where it reaches the maximum at 0.5.
- `delta` The smaller the change you wish to detect, the larger the sample size you'll need.
- `power` The higher the value means that that you wish to increase the confidence that you have in the result. Thus it means that you need a larger sample size.
- `sig_level` The smaller the value means that you wish to increase the confidence that you have in the result. Thus it means that you need a larger sample size.

# Analyze Quick Example's Result

Suppose you have run the test and you've obtain the total number of sample sizes and the total number of successes for both groups. Given these variables we can use it to calculate whether the proportional change was due to variation or not. To do so, we'll calculate the confidence interval for the difference, which can be done by calculating the difference plus and minus the z score times the standard deviation. Math Formula :
$$p_{experiment} - p_{control} \pm Z \sqrt{ p_{pooled}( 1-p_{pooled} )( \frac{1}{n_{experiment}} + \frac{1}{n_{control}} ) }$$

Where you can calculate the $p_{pooled}$, the pooled probability by the sum of the number of successes for both groups divided by the sum of the total number of sample size for both groups.

[`ABTest`][ABTest] Function that calculates the confidence interval. Input parameters :

- `count_control` The number of successes. This is equivalent to the number of people that clicked the button for the control group ( your original feature ).
- `sizes_control` The total number of sample size for the control group.
- The same notion can be applied to the experiment's variable `count_experiment` and `sizes_experiment`.
- Returns : A data.frame consisted of the mean and the confidence interval of the result.

```{r}

# parameters 
count_control 	 <- 974
sizes_control 	 <- 10072
count_experiment <- 1242
sizes_experiment <- 9886

ABTest <- function( count_control = count_control, 
					sizes_control = sizes_control,
					count_experiment = count_experiment, 
					sizes_experiment = sizes_experiment )
{
	# probability of each group 
	p_control <- count_control / sizes_control
	p_experiment <- count_experiment / sizes_experiment

	# @p : pooled probability
	# @std_error pooled standard deviation (error ) 
	p <- ( count_control + count_experiment ) / ( sizes_experiment + sizes_control )
	std_error <- sqrt( p * ( 1 - p ) * ( 1 / sizes_control + 1 / sizes_experiment ) )

	# 95 percent confidence interval's z score = 1.96, equivalent to 
	# qnorm( 0.975 )
	difference <- p_experiment - p_control
	confidence <- difference + c( -1, 1 ) * 1.96 * std_error

	return( data.frame( lower = confidence[1], 
						mean  = difference,
						upper = confidence[2] ) )
}

confidence <- ABTest( count_control = count_control, 
					  sizes_control = sizes_control,
					  count_experiment = count_experiment, 
					  sizes_experiment = sizes_experiment )
confidence

```

In order to launch a change, the change should be larger than the minimum detectable change that you wished to detect, or in other words it should be larger than your practical significance boundary. In our case, the value we've set was `r delta`. Base on the result above, we can denote that since even the lower bound of the confidence interval is larger than the value, we'll definitely launch the newer version of the click button.

Different scenarios of the output :

```{r, warning=FALSE,message=FALSE}

# fixed artifical plot
# using delta = 0.02 as the minimum detectable boundary 
library(ggplot2)

scenario <- as.character(2:6)
lower <- c( -0.008, 0.011, -0.025, -0.005, 0.015 )
mean  <- c( 0.005, 0.014, 0.005, 0.025, 0.025 )
upper <- c( 0.018, 0.017, 0.035, 0.055, 0.035 )

examples <- data.frame( scenario, lower, mean, upper )
examples <- rbind( cbind( scenario = "1", confidence ), examples )
examples$scenario <- factor( examples$scenario, levels = as.character(6:1) )

ggplot( examples, aes( mean, scenario, color = scenario ) ) + 
geom_point() + 
geom_errorbarh( aes( xmin = lower, xmax = upper ), height = 0.1 ) + 
geom_vline( xintercept = 0, color = "black" ) + 
geom_vline( xintercept = delta, color = "blue", linetype = "dotted" ) + 
geom_vline( xintercept = -delta, color = "blue", linetype = "dotted" ) +
scale_color_discrete( breaks = as.character(1:6) ) +  
labs( title = "Different Scenarios of Confidence Interval",
	  x = "confidence interval" )

```

1. **Scenario 1:** The case where even the lower bound of the confidence interval lies above the practical significance boundary. Accept the change of the new feature.
2. **Scenario 2:** The lower bound of the confidence interval lies below 0 and the upper bound lies below the practical significance boundary. There's no statistically significant change from 0 ( the confidence interval includes 0 ) and that you're also confident that there's not a practical significance change. Given this it's not worth the effort to launch the change.
3. **Scenario 3:** The lower bound of the confidence interval lies above 0 and the upper end lies below the practical significance boundary. This implies that you're confident that there is a positive change, but it's not practically significant. In other words, you're confident that there was a change, but you don't care about the magnitude of the change.
4. **Scenario 4:** Both the lower and upper bound of the confidence interval lies beyond the practical significance boundary. This means that the new feature could cause users to increase by the minimum detectable change or it could be cuasing them to decrease by the minimum detectable change.  
5. **Scenario 5:** The point estimate is beyond the practical significant line, the lower bound of the confidence interval, however, overlaps 0. This means that this change is in fact the effect that you care about, but there's also a chance that there might not be a change at all.
6. **Scenario 6:** The point estimate is beyond the practical significant line and the lower bound of the confidence interval is greater than 0. This is a situation that indicates the change has a chance of being practically significant and not being practically significant.

For the last three scenario, scenario 4 - 6: If your confidence interval includes your practical significance boundary, would you be sure that the change should not be launched? After all, it's reasonably likely that there was an effect you care about. In these cases, you should run an additional test with greater power if you have the time. But sometimes, you'll have to make a decision even though there's an uncertainty about how real your result is.

# Characterizing a Proper Metric

After walking through a simple A / B test example, we'll take a step back and see how to come up with a well-defined metric.

## Generate a high level metrices

For generating a high level metric, a good place to start is to think about the overall business objective. Below are a non-exhaustive list of techniques that might be able to help you come up with high level metrices that you might want to test out in your experiment.

1. **External Sources :** For these types of data, do not *match* them with internal data that you might already have at hand. Instead, look at a time series of both your
internally computed metric and the externally available one and see if the trends and seasonal variability line up.
2. **Gathering your own data :**
    - **User Experience Group :** With user experience group, you go really deep with a few users. It can take form of observing users doing tasks or you ask users to self-document their behaviors. After collecting the feedback it's usually a good idea to follow it up with a retrospective analysis. Meaning that if you observe something in a UER study, you could then go looking for that pattern in your data logs to see if that observation bears out and is worth creating a metric about.
    - **Focus Group :** This is basically where you recruit a bunch of users or potential users and bring them together for a group discussion or you can ask them questions. Though you run the risk of the answers get dominated by "loud voices".
    - **Survey :** The main advantage of surveys is that you can typically reach thousands of users. The disadvantages are that the data is self-reported and users don’t always give full or truthful answers.

    Qualitative methods such as User Experience Group and Focus Group are much better suited for answering questions about why or how to fix a problem. Whereas quantitative methods like Surveys do a much better job answering how many and how much types of questions. Having such numbers from the survey helps prioritize resources and allows you to focus on the biggest issues at hand.

## Turning it into a well-defined metrices

Recall that in the examples above, we've chosen the high-level metric click-through probability, which is defined by the unique number of people how click the button versus the number of users who visited the page that the button was located.

But to actually use implement the definition, we'll also have to address some other questions. Such as : If the same user visits the page once and comes back a week or two later, do we still only want to count that once? Usually, it is preferred that we count these separately, thus we'll also need to specify a time period.

So given that users are usually recorded by cookies online. One version of the fully-defined metric will be : For each minute ( choose any time interval that you prefer ), the number of cookies that clicked divided by the number of cookies that interacted with the page.

Another way of implementing this is to calculate : For each minute the number of pageviews ( assign each pageview with an id ) with a click divided by the number of pageviews.

## Segmenting and filtering your traffic

Creating segments and filtering out data can be useful for evaluating metric definitions, because, well, you can look at how the different definitions vary by segment.

For example, you're looking at your metric of total active users over time and you see a spike in one of the timelines. After confirming that this is not caused by seasonal variation, we can look at different segment of our visitors to see if one of the segment is causing the spike. Suppose we have chosen segment to be geographic, it might just happen that we've identify a large proportion of the traffic is generated by a specific region and it might be best for us to dig deeper and understand why.

To develop an understanding about your data and system, we can plot different metrices together, such as the rate ( calculated by the number of clicks divided by number of pageviews ) and probability by different segments to determine the differences in the output.

## Summarizing Metric

Apart from click-through probability, there're four other common summarizations that might be useful.

1. **sums and counts** : e.g. how many users visited the homepage.
2. **distributional metric** : e.g. quantiles. Mean age of users who completed the course ( though keed in mind that mean is sensitivity to outliers ).
3. **probabilities and rates** : Recall the subtle differences between the definition of probabilities and rates here. Probability means that given a time interval, whether a person clicked on the button or not ( takes the number of 0 / 1 ) divided by the total number users that interacted with the page but did not click the button during that period of time. As for rate, it refers to calculating the total number of clicks divided by the total number of pageviews.
4. **ratios.** e.g. the probability that a user clicks on a revenue generating click divided by any other click.

## Measuring Sensitivity and Robustness

It's the idea that you want to choose a metric that picks up the changes that you care about but robust against the changes that you don't care about, meaning that it doesn't move a lot when nothing interesting happens. It's an important thing to keep in mind when choosing the distributional metric as your summary metric.

e.g. When testing out things like video latency, you plot different summary metric of comparable video's latency, such as the mean, 90 percentile. In theory there shouldn't be that much of a difference between these videos, if there is, then it means that the summary metric isn't robust enough. And suppose you increase the video quality across these comparable videos and the mean does not change at all, then this probably indicates that choosing the mean as the summary metric isn't sensitive enough.

# Designing an Experiment

Suppose you have chosen the suitable metrics to be used for the test, now it's time to apply that and work through the decisions you need to make when you're actually designing an experiment. This can be boiled down to three main part.

1. Define what you mean as a "subject" in your experiment and control. In other words, what are the units in the population that you're going to be running the test on and comparing? 
2. Choose the "populations". e.g. Everyone or only subjects in the U.S. ? You need to be sure that you're computing the metrices on equivalent populations ( This holds true for subject ).
3. Determine the size of the experiment. This will also affect the duration of the experiment.

## Define the Subject

What you need to do is to decide how to assign events to either the control or the experiment. You can randomly assign every event, but if you have a user-visible change then this might not be a good idea. Since what will happen is : The user will see the change when they first visit the page and after reloading the page, the user will not see the change (leading to confusion). So for when conducting user-visible change, you want to assign people as oppose to events.

The next question is how do we define a user? There're three commonly used categories, namely user id, anonymous id (cookie) and event.

1. **user id :** e.g. Log in user names. Choosing this as the proxy for your user means that all the events that corresponds to the same user id are either in the control or experiment group, regardless of whether that user is switching between a mobile phone or desktop. This also means that if the user has not log in then he / she will neither be assgined to a control or experiment group.

2. **anonymous id (cookie) :** The cookie is specific for a browser and device, thus if the user switches from Chrome to Firefox, they'll be assigned to a different cookie. Also note that users can clear the cookie, in which case the next time they visit the website they'll get assigned to a new cookie even if they're still using the same browser and device. For experiments that will be crossing the sign-in border, using cookie is preferred. e.g. Suppose you're changing the layout of the page or locations of the sign in bar then you should use a cookie.

3. **event :** Should only be used when you're testing a non-user-visible change. e.g. page load time.

## Define the Population

If you think you can identify which population of your users will be affected by your experiment, you might want to target your experiment to that traffic ( e.g. changing features specific to one language's users ) so that the rest of the population won't dilute the effect.

Next, depending on the problem you're looking at, you might want to use a **cohort** instead of a population. A cohort makes much more sense than looking at the entire population when testing out learning effects, examining user retention or anything else that requires the users to be established for some reason.

A quick note on **cohort**. The gist of cohort analysis is basically putting your customers into buckets so you can track their behaviours over a period of time. The term cohort stands for a group of customers grouped by the timeline ( can be week, month ) where they first made a purchase ( can be a different action that's valuable to the business ). In a cohort, you can have roughly the same parameters in your two user group, which makes them more comparable.

For example, you're an educational platform has an existing course that's already up and running. Some of the students have completed the course, some of them are midway through and there're students who have not yet started. If you want to change the structure of of one of the lessons to see if it improves the completion rate of the entire course and they started the experiment at time X. For students who have started before the experiment initiated they may have already finished the lesson already leading to the fact that they may not even see the change. So taking the whole population of students and running the experiment on them isn't what you want. Instead, you want to segment out the cohort, the group of customers, that started the lesson are the experiment was launched and split that into an experiment and control group.

## Size and Duration

You can size your experiment based on your significance level, power and minimum detectable change like we've mentioned in the example. But you can also decrease the size of your experiment by targeting experiment to specific traffic. Because you're only looking at a subset of your traffic, you might need a bigger change before it to matters to the business ( Recall that the smaller the change you wish to detect, the larger the sample size you’ll need ). On other note although this might not reduce the time frame of the experiment, but this allows other experiments to be run on the same time.

After determining the size of the experiment it's time to translate that into a set of practical decisions including :

1. When do I want to run the experiment.
2. What fraction of the traffic are you going to send through the experiment. This will be closely tied to the duration of the experiment. There're some reasons that you might not want to run the experiment on all of your traffic to get the result faster. 
    - The first consideration might be you're just uncertained of how your users will react to the new feature, so you might want to test it out a bit before you get users blogging about it. To same notion applies to riskier cases, such as you're completely switching your backend system, if it doesn't work well, then the site might go down. 
    - Another consideration is that you may want to run the test a bit longer to make sure that the experiment captures the different user behaviour of weekdays, weekends.
    - The final one is that you may be running multiple tests at the same time and if you wish to make them directly comparable ( different level or attributes of the same feature ), then the easiest way is to run them at the same time. So if there's a holiday during the timespan that you're running the experiment, then all of the experiments should be equally affected.

# Analyzing Results

## Sanity Checks

Before jumping into analyze the result, we should check that the population size metric and any other metric that you don't expect to change shouldn't change while you're running the experiment, or in a sense that they're invariant and shouldn't differ between the experiment and control group.

Checking invariants example :

For instance, after running your experiment for a week, you've discovered that the total number of users assigned to the control group is 64454 and the total number of users assigned to the experiment group 61818. How would you figure out whether the difference is within expectation given that each user is randomly assigned to the control or experiment group with a probability of 0.5 ?

This is equivalent to saying out of a total 126272 ( 64454 + 61818 ) users, is it surprising to see if 64454 users are assigned to the control group, which is essentially a binomial distribution. Knowing this information, we can construct a confidence interval to test if the number lies within the confidence interval. The confidence interval can be calculated by the mean plus and minus the z-score times the standard error.

$$ mean \pm Z \sqrt{ np(1-p) } $$

Where the mean is expected number of users in the control / experiment group, which is simply the total number of the two groups times p (0.5). And the standard error of a binomial distribution is $\sqrt{ np(1-p) }$.

```{r}

group1 <- 64454
group2 <- 61818
SanityCheck <- function( group1, group2 )
{
	n <- group1 + group2
	confidence <- n * 0.5 + c( -1, 1 ) * 1.96 * sqrt( n * 0.5 * 0.5 ) 
	return( confidence )
}
( sanity <- SanityCheck( group1, group2 ) )

```

The result shows that 64454 does not lie within the range of the computed confidence interval and therefore it indicates that cookies may not be split equally.

When this kind of situation happens it's usually best to go back to the day by day data to get a better idea of what could be going wrong. One good thing is to check whether any particular day stands out, or it is just an overall pattern. If it is an overall pattern, then it is suggested that we should check if something went wrong with the experiment setup before proceeding on to analyzing the result.

## Analyzing the Result

One thing to keep in mind is that the goal of analyzing the result is not only about looking for statistical significance , but also about making a business decision about whether your experiment has favorably impacted your metric. So if you have an experiment that statistically improves the experience for thirty percent of the users but is neutral for everyone else, do you want to launch it as is, or do you want to try and make it better?

**Some other takeaways: **

A / B testing is an iterative process, some common mistakes that you want to avoid is :

1. Falling into the trap of “We already tried that”. A hypothesis can be implemented in so many different ways. If you just do one headline test and say "we tried that," you’re really selling yourself short.
2. Not testing continually or not retesting after months or years. Just because you tested a variation in the past doesn’t necessarily mean that those results are going to be valid a year or two from now.
3. Optimizing for the top of the funnel, rather than the product. Understanding what the customers want so that you can make the product better. Ultimately, you can’t simply test your headlines and get people to like your product more.

**Sources: **

Source [code](https://github.com/ethen8181/Business-Analytics/blob/master/3_AB_test/AB_test.R) to the documentation.

Bugs or comments can be filed [here](https://github.com/ethen8181/Business-Analytics/issues).

# Reference

1. Determining the sample size for A / B test : https://signalvnoise.com/posts/3004-ab-testing-tech-note-determining-sample-size
2. Z-score distribution table : http://www.utdallas.edu/dept/abp/zscoretable.pdf
3. Additional techniques to generating metrices : https://storage.googleapis.com/supplemental_media/udacityu/3954679115/additional_techniques.pdf
4. Pitfalls of A / B test : http://www.forbes.com/sites/sujanpatel/2015/10/29/how-to-do-ab-testing-right-and-avoid-the-most-common-mistakes-marketers-make/#2715e4857a0b54f535a44349

# R Session Information

```{r}

sessionInfo()

```

[ABTest]: https://github.com/ethen8181/Business-Analytics/blob/master/3_AB_test/AB_test.R