### 2016-04-21 articles/articles.Rmd
bayes done for now

---
title: "Articles"
date: "`r Sys.Date()`"
author: "Ming-Yu Liu"
output:
  rmdformats::readthedown:
    highlight: pygments
---

# Interviews and Careers

- Land any job you want. [[blog](http://landanyjobyouwant.com/)]
- The career advice no one tells you. [[Quartz](http://qz.com/640112/the-career-advice-no-one-tells-you/)]
- Wanted in College Graduates: Tolerance for Ambiguity. [[Linkedin Pulse](https://www.linkedin.com/pulse/wanted-college-graduates-tolerance-ambiguity-jeff-selingo)]
- How to set goals: OKRs. [[Youtube](https://www.youtube.com/watch?v=mJB83EZtAjc)]
- Meetings that don’t suck. [[Youtube](https://www.youtube.com/watch?v=rU8nv38E0Qk)]

# Analytics

- Using Machine Learning to Predict Customer Behaviour [[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/using-machine-learning-to-predict-customer-behaviour)]
    - Used with Complaints Management, Customer Upsell and Customer Retention. Questions include How to anticipate and manage the dissatisfaction of the customers?', 'How to sell more and better to existing customers?' and 'How to keep the most profitable customers at an appropriate cost?'
    - Action: You identify a target behaviour, in our case, it is a specific customer action, like filing a complaint/not filing a complaint, accepting or rejecting an upsell offer, cancelling / not cancelling a service. Then perform supervised learning
- Doing Data Science Right — Your Most Common Questions Answered. [[Review](http://firstround.com/review/doing-data-science-right-your-most-common-questions-answered/?imm_mid=0e29cf&cmp=em-data-na-na-newsltr_20160413)]
- Principles of good data analytics. [[blog](http://www.gregreda.com/2014/03/23/principles-of-good-data-analysis/)] 
- Why I Ditched My Fitbit and What This Means for Analytics. [[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/why-i-ditched-my-fitbit-and-what-this-means-for-analytics)]
- Know the Difference between Customer Service and Customer Experience. [[Harvard Business Review](https://hbr.org/sponsored/2016/03/know-the-difference-between-customer-service-and-customer-experience?utm_source=twitter&utm_medium=social&utm_campaign=harvardbiz)]
- Two articles about saying no to p-values. [[R-bloggers](http://www.r-bloggers.com/after-150-years-the-asa-says-no-to-p-values/)][[R-bloggers](http://www.r-bloggers.com/its-not-the-p-values-fault-reflections-on-the-recent-asa-statement-relevant-r-resources/)]
- The 9 Best Marketing And Sales Analytics You Should Know About.[[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/the-9-best-marketing-and-sales-analytics-you-should-know-about)]
- Cohort analysis. [[Youtube](https://www.youtube.com/watch?v=NyhVdGmnh0I)]
- Ted Talk : How to use data to make a hit TV show. [[Ted](http://www.ted.com/talks/sebastian_wernicke_how_to_use_data_to_make_a_hit_tv_show?utm_campaign=social&utm_medium=referral&utm_source=facebook.com&utm_content=talk&utm_term=technology)]

# Business

- Facebook is the new Excel. [[blog](http://alexmuir.com/facebook-is-the-new-excel)]
- We need companies to start innovating instead of imitating success stories. [[Hustle](http://thehustle.co/snapchat-vs-spacex-battle-of-the-imitators-and-innovators)]
- My 4+ Years At Uber Taught Me These Key Lessons. 1. May the supply be with you ; 2. Don't listen, show them. [[Linkedin](https://www.linkedin.com/pulse/three-things-i-learned-during-4-years-uber-michael-pao)]

# Learn Data Science

- The MBA Data Science Toolkit: 8 resources to go from the spreadsheet to the command line. [[Medium](https://medium.com/@dmca/the-mba-data-science-toolkit-8-resources-to-go-from-the-spreadsheet-to-the-command-line-cbb59ea82144#.pjph91n3m)]