### 2017-01-14 R/data_table/data_table.Rmd

---
title: "R's data.table"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

> The **source of this tutorial**, including the example datasets, is available [here on GitHub](https://github.com/ethen8181/Business-Analytics/blob/master/R/data_table)

# SQL Joins

For joining data.tables, the basic syntax are:

JOIN type   | DT syntax         | data.table::merge() syntax
----------- | ----------------- | --------------------------
INNER       | X[Y, nomatch = 0] | merge(X, Y, all = FALSE)
LEFT OUTER  | Y[X]              | merge(X, Y, all.x = TRUE)
RIGHT OUTER | X[Y]              | merge(X, Y, all.y = TRUE)
FULL OUTER  | -                 | merge(X, Y, all = TRUE)

---

The example data consists of the following two tables:

```{r}

library(data.table)
setwd("/Users/ethen/Business-Analytics/R/data_table")
employees <- fread("data/employees.csv")
departments <- fread("data/departments.csv")
knitr::kable(employees, caption = "Table Employees")
knitr::kable(departments, caption = "Table Departments")

```

## Inner Join

*Figure linked from [Pinal Dave][Pinal]*  
![Figure1 from Pinal Dave](http://www.pinaldave.com/bimg/March09UG/inner%20join.jpg)

**the *INNER JOIN* returns the rows with a match in both tables**

```{r}

# inner join with merge
merge(employees, departments, by = "Department")

# inner join with keys
# set the keys of the tables to represent the by:
setkey(employees, Department)
setkey(departments, Department)
# equivalent to setkeyv( departments, "Department" )

# note that you only need to set the keys once
# you can confirm if it works with
key(employees)

# perform the join,
# nomatch 0 means no rows will be returned 
# from not matched rows from the Right table
employees[ departments, nomatch = 0 ]

```

As we'll see later, without anything else, `TABLE_X[TABLE_Y]` returns a **right outer join**; setting `nomatch=0` returns a **inner join**.

**note:**

> Employee #6, Ferdinand, has not been returned, as he's in a yet to be created department 21.
> Neither has department #14, Research, as there are no employees.

## Left Outer Join

*Figure linked from [Pinal Dave][Pinal]*  
![Figure2 from Pinal Dave](http://www.pinaldave.com/bimg/March09UG/left%20join.jpg)

**the *LEFT OUTER JOIN* returns all the rows from the left table, filling in matched columns (or NA) from the right table**

```{r}

# left join with merge, notice that the column's ordering is different
merge(employees, departments, by = "Department", all.x = TRUE)
departments[employees]

```

**note:**

> Employee #6, Ferdinand, has been returned with department name as NA.  
> Department #14, Research, has not been returned.

## Right Outer Join

*Figure linked from [Pinal Dave][Pinal]*  
![Figure2 from Pinal Dave](http://www.pinaldave.com/bimg/March09UG/right%20join.jpg)

**the *RIGHT OUTER JOIN* returns all the rows from the right table, filling in matched columns (or else NA) from the left table**

```{r}

# right join with merge
merge(employees, departments, by = "Department", all.y = TRUE)
employees[departments]

# we can also do a not matched with !
# returning the rows in the employees table that are not in the
# department table
employees[!departments]

```

## Full Outer Join

*Figure linked from [Pinal Dave][Pinal]*  
![Figure2 from Pinal Dave](http://www.pinaldave.com/bimg/March09UG/outer%20join.jpg)

**the *FULL OUTER JOIN* returns all the rows from both tables, filling in matched columns (or else NA)**

```{r}

merge(employees, departments, by = "Department", all = TRUE)

```

# Other Tips and Tricks

## `.SD` and `.N`

First `.SD` refers to the subset of data for each group. The example below gets the sum of y and z and the number of rows (using `.N`) in each group while grouped by x.

```{r}

DT <- data.table( x = sample(1:2, 7, replace = TRUE), 
				  y = sample(1:13, 7), 
				  z = sample(1:14, 7) ) 
DT[ , c( lapply(.SD, sum), .N ), by = x ]
```

## Key Indexing

Set the keys and select the matching rows without using `==`.

```{r}

DT <- data.table( A = letters[ c(2, 1, 2, 3, 1, 2, 3) ],
				  B = c(5, 4, 1, 9, 8, 8, 6), C = 6:12 )
setkey(DT, A, B)

# matches rows where A == "b"
DT["b"]

# matches rows where A == "b" or A == "c"
DT[ c("b", "c"), ]

```

Since we did not provide any values for the second key column "B", it just matches c( "b", "c" ) against the first key column "A" and returns all the matched rows. Consider the example below that matches both keys.

```{r}

DT[ .("b", 5), ]

```

"b" is first matched against the first key column "A". And within those matching rows, 5 is matched against the second key column "B" to obtain row indices where both "A" and "B" match the given values.

```{r}

# Subset all rows where just the second key column matches
DT[ .(unique(A), 5), ]

```

## Selecting Parts of a Filtered Group

Select the first row of the `b` and `c` groups using `mult`.

```{r}

DT[ c("b", "c"), mult = "first" ]

```

When i is a data.table, `DT[ i, j, by = .EACHI ]` evaluates j for the groups of DT that each row in i joins to. We call this grouping by each i. See difference with and without `.EACHI` below.

```{r}

# we can use [] for .SD to subset the data
DT[ c("b", "c"), .SD[ c(1, .N), ] ]
DT[ c("b", "c"), .SD[ c(1, .N), ], by = .EACHI ]

```

Notice that `.N` here is used to retrieve the last row.

## R Session Information

```{r}
sessionInfo()
```

## Reference

- [Joining data in R using data.table](https://rstudio-pubs-static.s3.amazonaws.com/52230_5ae0d25125b544caab32f75f0360e775.html)
- [Data analysis in R using data.table](https://rpubs.com/davoodastaraky/dataTable)

[Pinal]:http://blog.sqlauthority.com/2009/04/13/sql-server-introduction-to-joins-basic-of-joins/